Species richness destabilizes trophic communities in freshwater ecosystems

### 2020-02-19 manuscript/bef_stability/manuscript.Rmd
Add computation of cv over basin + DiagrammeR

The biomass stability was decomposed in synchrony and weighted average CV
components according to @thibaut_understanding_2013.

@@

---
# Species richness to community attributes 
---
The reveal the mechanisms behind the relationship between species richness and
biomass stability, causal links were set from species richness to other
community attributes. (4) We hypothesized that TSR increased ST because
increasing species richness because of assembly processes. Species rich
community are more resistant to invasion and thus produces more ST
[@tilman_niche_2004; @shurin_how_2007].

```{r}
# Compute SEMs
stab_sem_rich_beta <- compute_stab_sem_rich_beta(.data = sem_data,
  random_effect = as.formula("~1|basin"))
stab_sem_meta <- compute_stab_sem_meta(.data = sem_data,
  random_effect = as.formula("~1|basin"))
stab_sem <- compute_stab_sem(.data = sem_data,
  random_effect = as.formula("~1|basin"))

bm_sem <- compute_prod_sem(.data = sem_data,
  random_effect = as.formula("~1|basin"))
bm_sem_rich_beta <- compute_prod_sem_rich_beta(.data = sem_data,
  random_effect = as.formula("~1|basin"))

library(ggraph)
p_stab_sem_rich_beta <- ggpiecewisesem(stab_sem_rich_beta)
p_stab_sem_meta <- ggpiecewisesem(stab_sem_meta, p_val_thd = 1)
p_stab_sem <- ggpiecewisesem(stab_sem)
p_bm_sem_rich_beta <- ggpiecewisesem(bm_sem_rich_beta)
p_bm_sem <- ggpiecewisesem(bm_sem)
plot_grid(
  p_stab_sem,
  p_stab_sem_rich_beta,
  p_bm_sem,
  p_bm_sem_rich_beta
)
```

@@

p_stab <- plot_grid(p_stab_div, p_stab_div_troph, ncol = 2, rel_widths = c(.4, .6))
```

@@

p_bm <- plot_grid(p_bm_div, p_bm_div_troph, ncol = 2, rel_widths = c(.4, .6))
```

@@

- Stability:
  - decreases globally with total species richness (TSR)
    - An increase of 1% of TSR decreased biomass stability by
      `r s_co_r`% (`r s_ci_r`)
  - was `r s_co_t`% (`r s_ci_t`) higher for higher trophic group
  - increased by `r s_co_r_t2`% (`r s_ci_r_t2`) and decreased by `r s_co_r_t3`
    (`r s_ci_r_t3`) with 1% increase in TSR for the lower and higher trophic
    group respectively

- Biomass:
  - increases with TSR 
    - An increase of 1% of TSR increased total biomass by
      `r b_co_r` % `r b_ci_r`
  - was `r b_co_t`% `r b_ci_t` higher for higher trophic group
  - increased by `r b_co_r_t2`% `r b_ci_r_t2` and decreased by `r b_co_r_t3`
    (`r b_ci_r_t3`) with 1% increase in TSR for the lower and higher trophic
    group respectively

@@

```{r sem-plot, fig.dim = c(10,10)}
#https://tex.stackexchange.com/questions/20496/drawing-different-tikz-shapes-parameterized-by-data-from-a-file?rq=1
#https://tex.stackexchange.com/questions/83888/how-to-plot-data-from-a-csv-file-using-tikz-and-csvsimple
to_plot <- c(1, 6, 7)
plot_grid(
  plot_grid(plotlist = map(p_l_rot[to_plot], ~.x$rotated),
      ncol = length(to_plot), labels = LETTERS[1]),
  plot_grid(p_stab_sem_rich_beta, p_bm_sem_rich_beta, labels = LETTERS[2:3]),
  ncol = 1,
  rel_heights = c(.4, .6)
)
```

@@

@@

### SEM linear relationship

@@

mod_stab_list <-
  list(
    nlme::lme(log_rich_tot ~ RC1 + RC2 + RC3 + RC4 + RC5,  random = ~ 1 | basin, data = sem_data),
    nlme::lme(piel ~ RC1 + RC2 + RC3 + RC4 + RC5 + log_rich_tot, random = ~ 1 | basin, data = sem_data),
    nlme::lme(ct ~ RC1 + RC2 + RC3 + RC4 + RC5 + log_rich_tot, random = ~ 1 | basin, data = sem_data),
    nlme::lme(t_lvl ~ RC1 + RC2 + RC3 + RC4 + RC5 + log_rich_tot, random = ~ 1 | basin, data = sem_data),
    nlme::lme(beta_bin ~ RC1 + RC2 + RC3 + RC4 + RC5 + log_rich_tot, random = ~ 1 | basin, data = sem_data),
    nlme::lme(log_sync ~ log_rich_tot + piel + ct + t_lvl + RC1 + RC2 + RC3 + RC4 + RC5,
      random = ~ 1 | basin, data = sem_data),
    nlme::lme(log_cv_sp ~ log_rich_tot + piel + ct + t_lvl
      + RC1 + RC2 + RC3 + RC4 + RC5, random = ~ 1 | basin, data = sem_data),
    lm(log_stab ~ log_cv_sp + log_sync, data = sem_data)
  )
names(mod_stab_list) <- c("richness", "pielou", "connectance", "weighted_mean_trophic_lvl", "beta", "synchrony", "cv_sp", "stability") 
colin_stab_sem <- map(mod_stab_list, car::vif)

@@

#formulas <- list(
  #"log_rich ~ RC1 + RC2 + RC3 + RC4 + RC5",
  #"piel ~ RC1 + RC2 + RC3 + RC4 + RC5"
#)
#test <- map(formulas, function(x, y, random_effect) {
  #lmer(as.formula(x), random= as.formula(random_effect), data = y)

#}, y =sem_data, random_effect = "~ 1 |basin")

@@

```{r, eval = FALSE}
mod_tlvl <- nlme::lme(t_lvl ~ scale(RC1) + scale(RC2) + scale(RC3) + scale(RC4) + scale(RC5) + log_rich_tot, random = ~ 1 | basin, data = sem_data)
plot(mod_tlvl)
sem_data %>%
  ggplot(aes(x = log10(richness_tot), y = t_lvl)) +
  geom_point()+
  geom_smooth(method = "lm")

sem_data %>%
  ggplot(aes(x = RC1, y = t_lvl)) +
  geom_point()+
  geom_smooth(method = "lm")
plot(nlme::lme(t_lvl ~ I(log10(RC1 + abs(min(RC1)))) + log_rich_tot, random = ~ 1 | basin, data = sem_data))
nlme::lme(t_lvl ~ log_rich_tot, random = ~ 1 | basin, data = sem_data)
semd_test <- sem_data %>%
  mutate(RC1b = RC1 + abs(min(RC1)))
ricker.mixed.fit <- nls(
  t_lvl ~ (Vm*(RC1 + b) / (K + (RC1 + b)) + d),
 #fixed = Vm + K + b + d ~ 1,
 #random = Vm + K + b + d ~ 1|basin,
 start= c(Vm=1, K = 0.1, b = .5, d = 2.5),
 data = semd_test,
 control = lmeControl(maxIter = 1e8, msMaxIter = 1e8) 
)
summary(ricker.mixed.fit)
x <- seq(min(sem_data$RC1), max(sem_data$RC1), length.out = 100)
sem_data$test_pred <- predict(ricker.mixed.fit)

@@

>>>>

The biomass stability was decomposed in synchrony and weighted average CV
components according to @thibaut_understanding_2013:

$$
CV_{com} = \overline{CV_{sp}} \times \sqrt{\phi}
$$

with $CV_{com}$ being the Coefficient of Variation of the community,
$\overline{CV_{sp}}$ the average coefficient of variation of population
pondered by biomass. $\phi$ is the synchrony, which is the ratio the total
community variance ($\sigma^2_{x_T}$) and the sum of population variance
($\sigma_{x_i}$) [@loreau_species_2008]:

$$
\phi = \frac{\sigma^2_{x_T}}{\sum_i{\sigma_{x_i}}}
$$

@@

---
# Species richness to community attributes 
---
To reveal the mechanisms behind the relationship between species richness and
biomass stability, causal links were set from species richness to other
community attributes. (4) We hypothesized that TSR increased ST because
increasing species richness because of assembly processes. Species rich
communities are more resistant to invasion and thus increase ST
[@tilman_niche_2004; @shurin_how_2007]. (5) We hypothesized also that TSR
decreased connectance following results from @thebault_stability_2010;
@winemiller_must_1989. But theoretically, some ecological networks can show the
inverse relationship [@winemiller_must_1989]. (6) We hypothesized also that TSR
will modulate average weighted trophic level because of food web assembly
processes. If *TSR increased and that new species are placed at random in the
network, we expected a positive relationship between Tlvl and TSR because
individuals of the higher trophic level have a higher body mass [to think
again]*. Finally, we set up causal relationships from community attributes to
stability components.

---
#  Community attributes to stability
---
We hypothesized that (7) TSR decreased synchrony between species by portfolio
effects [REF] and increased $\overline{CV_{sp}}$ [@tilman_biodiversity_2006] *to
complete*. We hypothesized that (8) ST increased $\overline{CV_{sp}}$ and that it
decreased synchrony. (9a) We expected that Tlvl had a positive or negative
relationship with synchrony, bottom-up effects should result in positive whereas
top-down in negative relationship. We expected a (9b) negative relationship
between Tlvl and $\overline{CV_{sp}}$ if predators stabilised population
fluctuations or if the increasing part of biomass in the predators decreased
$\overline{CV_{sp}}$ because of their higher body mass (if mean increase faster than
variance). (10a) Connectance was expected to decrease or increase $\overline{CV_{sp}}$
and (10b) synchrony because of top down effects or bottom up effects.

```{r}
# Compute SEMs
stab_sem_rich_beta <- compute_stab_sem_rich_beta(.data = sem_data,
  random_effect = as.formula("~1|basin"))
stab_sem_rich <- compute_stab_sem_rich(.data = sem_data,
  random_effect = as.formula("~1|basin"))
stab_sem_meta <- compute_stab_sem_meta(.data = sem_data,
  random_effect = as.formula("~1|basin"))
stab_sem <- compute_stab_sem(.data = sem_data,
  random_effect = as.formula("~1|basin"))

bm_sem <- compute_prod_sem(.data = sem_data,
  random_effect = as.formula("~1|basin"))
bm_sem_rich <- compute_prod_sem_rich(.data = sem_data,
  random_effect = as.formula("~1|basin"))
bm_sem_rich_beta <- compute_prod_sem_rich_beta(.data = sem_data,
  random_effect = as.formula("~1|basin"))

library(ggraph)
p_stab_sem <- ggpiecewisesem(stab_sem)
p_stab_sem_rich_beta <- ggpiecewisesem(stab_sem_rich_beta)
p_stab_sem_rich <- ggpiecewisesem(stab_sem_rich)

library(DiagrammeR)
graph <- build_diag_from_sem(stab_sem_rich, p_val_thd = 0.05)

p_bm_sem <- ggpiecewisesem(bm_sem)
p_bm_sem_rich <- ggpiecewisesem(bm_sem_rich)
```

@@

p_stab <- plot_grid(p_stab_div, p_stab_div_troph,
  ncol = 2, rel_widths = c(.4, .6),
  labels = c("A", "B") 
)
```

@@

p_bm <- plot_grid(p_bm_div, p_bm_div_troph, ncol = 2, rel_widths = c(.4, .6),
  labels = c("C", "D"))
```

@@

---
# Total biomass versus TSR
---
TSR increased total biomass (`r b_co_r`% `r b_ci_r` by 1% TSR increase). Total
biomass was `r b_co_t`% `r b_ci_t` higher in the high than in the low trophic group. However,
the increase of total biomass with the increase of TSR was lower in the high
than the low trophic group (`r b_co_r_t3`% `r b_ci_r_t3` and `r b_co_r_t2`%
`rb_ci_r_t2` respectively,  by 1% TSR increase).
---
# Stability versus TSR
---
However, biomass stability decreased with TSR (`r s_co_r`% `r s_ci_r` by 1%
TSR increase). Biomass stability was overall `r s_co_t`% `r s_ci_t` higher for the high
than the low trophic group. But biomass stability decreased by `r s_co_r_t3`% by
1% increase in TSR in the high trophic group and increased by `r s_co_r_t2`% `r s_ci_r_t2`
in the low trophic group.

@@

```{r sem-plot, fig.dim = c(10,10)}
#https://tex.stackexchange.com/questions/20496/drawing-different-tikz-shapes-parameterized-by-data-from-a-file?rq=1
#https://tex.stackexchange.com/questions/83888/how-to-plot-data-from-a-csv-file-using-tikz-and-csvsimple
to_plot <- c(1, 6, 7)
plot_grid(
  plot_grid(plotlist = map(p_l_rot[to_plot], ~.x$rotated),
      ncol = length(to_plot), labels = LETTERS[1]),
  plot_grid(p_stab_sem_rich, p_bm_sem_rich, labels = LETTERS[2:3]),
  ncol = 1,
  rel_heights = c(.4, .6)
)
```

@@

- R diagramm:
  - https://rstudio-pubs-static.s3.amazonaws.com/78926_5aa94ae32fae49f3a384ce885744ef4a.html

@@

### SEM linear relationships

```{r hist-sem-data}
sem_data %>%
  select(prod, log_rich_tot, log_sync, log_stab, log_cv_sp, RC1, RC2, RC3, RC4,
    RC5, t_lvl, ct, log_bm) %>%
  gather(variable, value) %>%
  ggplot(aes(x = value)) +
  geom_histogram() +
  facet_wrap(~variable, scales = "free")

@@

mod_stab_list <-
  list(
    nlme::lme(log_rich_tot ~ log_RC1 + log_RC2 + log_RC3 + RC4 + RC5,  random = ~ 1 | basin, data = sem_data),
    #nlme::lme(piel ~ log_RC1 + log_RC2 + log_RC3 + RC4 + RC5 + log_rich_tot, random = ~ 1 | basin, data = sem_data),
    nlme::lme(ct ~ log_RC1 + log_RC2 + log_RC3 + RC4 + RC5 + log_rich_tot, random = ~ 1 | basin, data = sem_data),
    nlme::lme(t_lvl ~ log_RC1 + log_RC2 + log_RC3 + RC4 + RC5 + log_rich_tot, random = ~ 1 | basin, data = sem_data),
    #nlme::lme(beta_bin ~ log_RC1 + log_RC2 + log_RC3 + RC4 + RC5 + log_rich_tot, random = ~ 1 | basin, data = sem_data),
    nlme::lme(log_sync ~ log_rich_tot + ct + t_lvl + log_RC1 + log_RC2 + log_RC3 + RC4 + RC5,
      random = ~ 1 | basin, data = sem_data),
    nlme::lme(log_cv_sp ~ log_rich_tot  + ct + t_lvl
      + RC1 + log_RC2 + log_RC3 + RC4 + RC5, random = ~ 1 | basin, data = sem_data),
    lm(log_stab ~ log_cv_sp + log_sync, data = sem_data)
  )
names(mod_stab_list) <- c("richness", "connectance", "weighted_mean_trophic_lvl", "synchrony", "cv_sp", "stability") 
colin_stab_sem <- map(mod_stab_list, car::vif)

@@

@@

```{r, eval = FALSE}
mod_tlvl <- nlme::lme(t_lvl ~ scale(RC1) + scale(RC2) + scale(RC3) + scale(RC4) + scale(RC5) + log_rich_tot, random = ~ 1 | basin, data = sem_data)
plot(mod_tlvl)
sem_data %>%
  ggplot(aes(x = log10(richness_tot), y = t_lvl)) +
  geom_point()+
  geom_smooth(method = "lm")
plot(x = log10(sem_data$RC1 + abs(min(sem_data$RC1))), y = sem_data$t_lvl)
sem_data %>%
  ggplot(aes(x = log10(RC1), y = t_lvl)) +
  geom_point()+
  geom_smooth(method = "lm")
plot(nlme::lme(t_lvl ~ I(log10(RC1 + abs(min(RC1)))) + log_rich_tot, random = ~ 1 | basin, data = sem_data))
nlme::lme(t_lvl ~ log_rich_tot, random = ~ 1 | basin, data = sem_data)
semd_test <- sem_data %>%
  mutate(RC1b = RC1 + abs(min(RC1)))
ricker.mixed.fit <- nls(
  t_lvl ~ (Vm*(RC1 + b) / (K + (RC1 + b)) + d),
 #fixed = Vm + K + b + d ~ 1,
 #random = Vm + K + b + d ~ 1|basin,
 start= c(Vm=1, K = 0.1, b = .5, d = 2.5),
 data = semd_test,
 control = lmeControl(maxIter = 1e8, msMaxIter = 1e8) 
)
summary(ricker.mixed.fit)
x <- seq(min(sem_data$RC1), max(sem_data$RC1), length.out = 100)
sem_data$test_pred <- predict(ricker.mixed.fit)

@@

```{r, eval = FALSE}
library(glmmTMB)
# Gaussian
mod_log_rich <- glmmTMB(
  log_rich_tot ~ log_RC1 + log_RC2 + log_RC3
  + RC4 + RC5 + (1 | basin), family = gaussian, data = sem_data)
summary(mod_log_rich)
plot(y = resid(mod_log_rich), x = fitted(mod_log_rich))

# Poisson 
mod_log_rich_poi <- glmmTMB(
  richness_tot ~ log_RC1 + log_RC2 + log_RC3
  + RC4 + RC5 + (1 | basin), family = poisson, data = sem_data)
summary(mod_log_rich_poi)
plot(y = resid(mod_log_rich_poi), x = fitted(mod_log_rich_poi))

# Truncate 
mod_log_rich_nbinom <- glmmTMB(
  richness_tot ~ log_RC1 + log_RC2 + log_RC3
  + RC4 + RC5 + (1 | basin), family = nbinom1, data = sem_data)
summary(mod_log_rich_nbinom)
plot(y = resid(mod_log_rich_nbinom), x = fitted(mod_log_rich_nbinom))
```

#### Total biomass

```{r}

mod_bm_list <-
  list(
    nlme::lme(log_rich_tot ~ RC1 + RC2 + RC3 + RC4 + RC5,  random = ~ 1 | basin, data = sem_data),
    #nlme::lme(piel ~ RC1 + RC2 + RC3 + RC4 + RC5 + log_rich_tot, random = ~ 1 | basin, data = sem_data),
    nlme::lme(ct ~ RC1 + RC2 + RC3 + RC4 + RC5 + log_rich_tot, random = ~ 1 | basin, data = sem_data),
    nlme::lme(t_lvl ~ RC1 + RC2 + RC3 + RC4 + RC5 + log_rich_tot, random = ~ 1 | basin, data = sem_data),
    #nlme::lme(beta_bin ~ RC1 + RC2 + RC3 + RC4 + RC5 + log_rich_tot, random = ~ 1 | basin, data = sem_data),
    nlme::lme(log_bm ~ log_rich_tot + ct + t_lvl + RC1 + RC2 + RC3 + RC4 + RC5,
      random = ~ 1 | basin, data = sem_data)
  )
names(mod_bm_list) <- c("richness", "connectance", "weighted_mean_trophic_lvl", "total_biomass") 
colin_bm_sem <- map(mod_bm_list, car::vif)

tmp <- map(colin_bm_sem, enframe)

# colinearity table
colin_bm_sem <- tibble(vif = tmp, mod = names(mod_bm_list)) %>%
  unnest(vif) %>%
  spread(name, value)

kable(colin_bm_sem)
```
